---
title: "Täuber et al. (2026) - Structural analysis"
author: "Alexander Täuber"
format:
  html:
    css: styles.css
    self-contained: true
    toc: true
    number-sections: true
project:
    execute-dir: project
execute:
  warning: false
  message: false
---

# Install the `coi` package
```{r}
#| label: Install coi
#| eval: false

# install.packages("remotes")
remotes::install_github("ataeub/coi", ref = "v2.1.0")
```

# Libraries & local scripts
```{r}
#| label: Preparations
#| eval: false

library(magrittr)

source(here::here("src/make_manifest.R"))
source(here::here("src/trees_to_tsps.R"))
source(here::here("src/compute_tsp_structure.R"))
source(here::here("src/compute_tsp_traits.R"))
```

# Set global variables

We can change the algorithm parameters for the run by changing the values in the parameters list.
```{r}
#| label: Global variables
#| eval: false

single_tree_p <- here::here("data/raw/single_trees")
tree_data_p <- here::here("data/processed/tree_level_data.csv")
output_p <- here::here("data/results/tsp_data.csv")

parameters <- list(
  vox_res = 0.01,
  coi_d_max = 0.645,
  cci_strata_size = 0.3,
  cci_strata_hull_type = "concave",
  boxdim_threshold = 0.1,
  filter_dist = 0.3
)
```

# Prepare the data for processing

- `make_manifest()` examines the directory with the single tree point clouds and compares it to our list of TSPs with pre-existing data on them (ID, species, etc.)
- `calculate_tsp_traits()` extracts the trait related data. In our case only the TSP-LTD and the species composition of the TSP.
- `trees_to_tsps()` is basically just widening our data from one-row-per-tree to one-row-per-TSP
Here we  and compare it 
```{r}
#| label: Data preparation
#| eval: false

tree_data <- make_manifest(
  tree_table_path = tree_data_p,
  single_tree_path = single_tree_p
)

traits <- tree_data %>%
  calculate_tsp_traits()

tsp_data <- tree_data %>%
  trees_to_tsps() %>%
  dplyr::filter(!is.na(tree1_pc_path) & !is.na(tree2_pc_path)) %>%
  dplyr::left_join(traits, by = "tsp")
```

# Structural data processing

- The magic happens in `compute_tsp_structure()` where the functions from the `coi` package are used to extract the structural variables from the point clouds
- Since the computation is quite resource intensive we use `future` and `furrr` to parallelize the computation to n-1 available physical CPU cores

```{r}
#| label: Data processing
#| eval: false

# Using the amount of physical cores minus one
cores_to_use <- parallelly::availableCores(logical = FALSE) - 1
future::plan(multisession, workers = cores_to_use)
progressr::handlers("cli")
options(progressr.show_after = 0)

progressr::with_progress({
  p <- progressr::progressor(steps = nrow(tsp_data))

  full_data <- tsp_data %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      results = furrr::future_pmap(
        list(tree1_pc_path, tree2_pc_path),
        ~ compute_tsp_structure(..1, ..2, parameters, p),
        .options = furrr::furrr_options(seed = TRUE)
      )
    ) %>%
    tidyr::unnest_wider(results)
})

readr::write_csv(full_data, output_p)
```

# COI - $d_{max}$ sensivity analysis

```{r}
#| label: Sensivity analysis
#| eval: false
dmax_values <- c(seq(0, 1, 0.2), 2, 4, 8, 16, 100)
vox_res <- 0.01

cores_to_use <- parallelly::availableCores(logical = FALSE) - 1
future::plan(multisession, workers = cores_to_use)
progressr::handlers("cli")
options(progressr.show_after = 0)

progressr::with_progress({
  p <- progressr::progressor(steps = nrow(tsp_data) * length(dmax_values))

  coi_sensitivity <- tsp_data %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      results = furrr::future_pmap(
        list(tree1_pc_path, tree2_pc_path),
        ~ compute_tsp_coi_sensitivity(..1, ..2, dmax_values, vox_res, p),
        .options = furrr::furrr_options(seed = TRUE)
      )
    ) %>%
    tidyr::unnest(results)
})

output_p <- here::here("data/results/coi_sensitivity.csv")
readr::write_csv(coi_sensitivity, output_p)
```